A large leading tail of atmospheric escape in the HAT-P-67 b system

### 2023-02-14 paper/paper1/ms.tex
working on the same document on two computers asynchronously... oops

If we attribute the TESS modulation signal to stellar activity, we obtain a rotation period $P_\mathrm{rot}$ between 4.7 and 5.9 days.  \citet{2017AJ....153..211Z} estimated $v\sin{i}=35.8\pm1.1$ km/s, or lower ($v\sin{i}=30\pm1.1$ km/s) if macroturbulence was accounted for.  We obtain a combined geometric constraint $2.85< R_\star \sin{i}  < 4.2 $ in units of $R_\odot$.  The lower end of these constraints is consistent with  the 1-$\sigma$ range of $2.0 < R_\star < 3.2$ provided in \citet{2017AJ....153..211Z} from different isochrones and orbit fits, and assuming a near-equator-on orbital inclination.  The Gaia \emph{DR3}-derived distance is 8\% greater than the \emph{Gaia} DR1 estimate used in \citet{2017AJ....153..211Z}; the revised larger distance would tend to increase the radius estimate by 16\% to match the observed SED in isochrone-fitting.  We would therefore expect an isochrone-derived $R_\star$ range closer to $2.5-3.14 R_\odot$, where we have dropped the 1-$\sigma$ uncertainties which had included the less certain Gaia DR 1 parallax and its uncertain correction terms.  For a fixed $R_p/R_\star$, the larger $R_\star$ implies a proportionally larger planet radius, revising the already-low density of HAT-P-67 even lower.

@@

Instead, the exteme mass loss and credibility of the ohmic diffusion model motivates the prospect of star-planet-interaction (SPI) that we consider now.  In this scenario, the magnetic field of the planet permeates the space between the star and planet.  Magnetic perturbations propagate via planetary magnetic fields with the Alfv\'en speed.  The planet can interact with the star if the Alfv\'en speed exceeds the stellar wind speed controling the bulk motion of intervening medium.  The magnetic field of the planet is not known, but the ohmic dissipation mechanism favors a limited range in the magnetic field strength: large enough to cause significant dissipative energy generation, but not too large to shut down motion entirely.  \citet{2011ApJ...738....1B} estimate this range as between about 1-30 Gauss.  With

A few observations disfavor this interpretation.  First, the non-detection of variability in the \ion{Ca}{2} H and K lines and H~$\alpha$ lines.

>>>>

If we attribute the TESS modulation signal to stellar activity, we obtain a rotation period $P_\mathrm{rot}$ between 4.7 and 5.9 days.  \citet{2017AJ....153..211Z} estimated $v\sin{i}=35.8\pm1.1$ km/s, or lower ($v\sin{i}=30\pm1.1$ km/s) if macroturbulence was accounted for.  Assuming a stellar radius of 2.1-2.7 $R_\odot$, we obtain combined geometric constraint $3.0 < \frac{P_\mathrm{rot}}{\sin{i}}  < 4.5 $ days.  This range is close to---but slightly lower than---the TESS modulation period.   The Gaia \emph{DR3}-derived distance is 8\% greater than the \emph{Gaia} DR1 estimate used in \citet{2017AJ....153..211Z}; the revised larger distance would tend to increase the radius estimate by 16\% to match the observed SED in isochrone-fitting.  We would therefore expect an isochrone-derived $R_\star$ range closer to $2.5-3.14 R_\odot$, where we have dropped the 1-$\sigma$ uncertainties which had included the less certain Gaia DR 1 parallax and its uncertain correction terms.  For a fixed $R_p/R_\star$, the larger $R_\star$ implies a proportionally larger planet radius, revising the already-low density of HAT-P-67 even lower.

@@

Instead, the exteme mass loss and credibility of the ohmic diffusion model motivates the prospect of star-planet-interaction (SPI) that we consider now.  In this scenario, the magnetic field of the planet permeates the space between the star and planet.  Magnetic perturbations propagate via planetary magnetic fields with the Alfv\'en speed.  The planet can interact with the star if the Alfv\'en speed exceeds the stellar wind speed controling the bulk motion of intervening medium.  The magnetic field of the planet is not known, but the ohmic dissipation mechanism favors a limited range in the magnetic field strength: large enough to cause significant dissipative energy generation, but not too large to shut down motion entirely.  \citet{2011ApJ...738....1B} estimate this range as between about 1-30 Gauss.  Such a modest planetary field appears adequate to maintain a magnetosphere \citep{2019ApJ...881..136S}.

The non-detection of variability in the \ion{Ca}{2} H and K lines and H~$\alpha$ lines appears to disfavor this SPMI interpretation.
